The Birth of a Galaxy: Population III Metal Enrichment and Population II Stellar Populations

### 2010-10-20 science_ms.tex
Revised version after another reading.

The first (Pop III) stars are metal-free and have a large
characteristic mass and suppressed fragmentation in its protostellar
collapse \cite{Abel00, ABN02, Bromm02_P3, Yoshida03, OShea07a}.  A
fraction of these stars enrich the surrounding intergalactic medium
(IGM) when they go supernova, which can happen in stars $\lsim 40~\Ms$
in Type II supernovae (SNe) or in stars roughly between 140 \Ms~and
260 \Ms~in pair-instability SNe (PISNe; \cite{2002ApJ...567..532H}).
The host halo and the neighboring halos are then enriched with this
ejecta.  There exists a critical metallicity that is $\sim 10^{-6}
Z_\odot$ if dust cooling is efficient \cite{Omukai05,
  Schneider06_Frag, clark08} and $\sim 10^{-3.5} Z_\odot$ otherwise
\cite{Bromm01, 2009ApJ...691..441S}, where the gas can cool rapidly,
lowering its Jeans mass.  An intermediate characteristic mass of
$\sim10~\Ms$ can be occur if the gas cooling is suppressed to the
cosmic microwave background (CMB) temperature \cite{Larson98,
  Tumlinson07_IMF, 2009ApJ...691..441S}.  The resulting Population II
star cluster will thus have a lower characteristic stellar mass than
its metal-free progenitors.  In the local universe, stars in the Milky
Way halo and dwarf spheroidal (dSph) galaxies may be connected to
these first stellar clusters.  In the halo of the Milky Way, the most
metal-poor stars have a metallicity floor of [Z/H] = --4
\cite{Beers05}.  In dwarf spheroidal (dSph) galaxies, this floor is a
factor of 10 higher \cite{Tolstoy04, Helmi06}, however several metal
poor stars with $-4 < [\mathrm{Z/H}] < -3$ have been recently
discovered in dSphs \cite{Tafelmeyer10, Frebel10_Obs}.

The transition from Pop III to Pop II star formation (SF) is solely
dependent on the propagation of metals from the SNe remnants into
future sites of SF.  Their flows are complex because of the
interactions between the SN blastwave, cosmological accretion and halo
mergers, and nearby stellar feedback.  In minihalos ($M_{\rm h} \sim
10^6~\Ms$), radiation from the massive Pop III drives a 30 \kms~shock,
which is 10 times greater than the escape velocity of the halo, and
leaves behind a warm ($3 \times 10^4$ K) and diffuse (0.5 \cubecm)
medium \cite{Kitayama04, Whalen04, Abel07}.  This aids in the
expansion of the blastwave because it delays the transition to the
Sedov-Taylor and snowplow phases.  In PISNe, approximately half of the
metals stay in the IGM with a metal bubble size of a few kpc.  The
remaining metals fall back into the host halo.  The blastwave may
induce SF in nearby halos through the compression of the gas
\cite{Ferrara98}; however timescales for metal mixing into the dense gas are
many dynamical times \cite{Cen08} for shock velocities $\lsim100~\kms$.

Numerical simulations are useful to detangle and study these
complexities and the transition from Pop III to II stars.  In this
Report, we present the simulations that include both types of SF, and
their radiative and mechanical feedback.  We utilize an initial mass
function (IMF) for individual Pop III stars, based on the latest
simulation results.  The methods used here incorporate and link
together recent results from metal-enriched and metal-free star
formation, the critical metallicity, and pair-instability supernovae.
Furthermore we spatially distinguish metal enrichment from Pop II and
Pop III stars.  This is the first time that all of these processes
have been interconnected in a simulation that resolves star-forming
minihalos to study the transition to Pop II star formation.

@@

Here we present the gaseous and stellar evolution of two selected
halos in the simulation: one that has an early mass buildup but no
major mergers after $z=12$, and one that experiences a series of major
mergers between $z=10$ and $z=7$.  We name the halos ``quiet'' and
``intense'', respectively.  We start with the mass accretion history
and enrichment of the halos.  We then compare the nature of the
stellar populations in these halos.

We illustrate the state of the simulation at $z=7$ in Figure
\ref{fig:projections} with density weighted projections of gas
density, temperature, and metallicity, showing the entire box and
focusing on the two halos of interest.  The quiet and intense halos
are located in the bottom center and bottom right of the full box,
respectively.  Radiative and mechanical feedback create a multi-phase
medium inside these halos, which are embedded in a warm and ionized
IGM.

@@

The gas fractions of both halos are depleted from 0.15 to 0.08 by
outflowing shocks driven by ionization fronts and blastwaves in their
initial starbursts.  The quiet halo never recovers to the cosmic
fraction $\Omega_b/\Omega_M$ for the following reason.  The lack of
major mergers of halos above the filtering mass leads to a small gas
fraction because halos with $M < M_f$ are likely photo-evaporated,
hosting diffuse warm gas reservoirs instead of cold dense cores.
During this quiescent period, the halo mainly accretes warm diffuse
gas from the filaments and IGM.  In contrast, the intense halo grows
from major mergers of halos with $M > M_f$.  The progenitor halos
involved in the major mergers are able to host molecular clouds and
have, in general, higher gas fractions.  Between $z=10$ and $z=8$, the
gas fraction increases from 0.07 to 0.12 until it jumps to 0.14 when a
gas-rich major merger occurs.  The stellar mass accordingly increases
with the ample supply of gas during this period.  The gas fraction
then decreases by 0.01 when galactic outflows are generated from the
starburst associated with this major merger.

\subsection*{Metallicity evolution}

The evolution of the stellar and gas metallicity of both halos are
illustrated in Figure \ref{fig:evo}b.  PISNe from Pop III stars enrich
the nearby IGM out to a radius of 10 kpc and provides a metallicity
floor of $[Z_3/H] \sim -3$, given $M_{\rm char} = 100 \Ms$.  The
metallicity from Pop II SNe initially enrich the ISM of both halos to
an average $[\bar{Z}_2/H]$ between --3.5 and --3.  The quiet halo is
enriched to $[\bar{Z}_2/H] \sim -2.5$ until the metal-poor inflows
balance the metal-rich galactic outflows, which are directed in the
polar directions of the gas disk, keeping the adjacent filaments
metal-poor.  These features and a well-mixed ISM
(cf.~\cite{Wise08_Gal, Greif10}) are apparent in the metallicity
projections in Figure \ref{fig:projections}.  The average stellar
metallicities is equal or slightly higher up to 0.5 dex than
$[\bar{Z}_2/H]$.

In the intense halo, the first few Pop II star clusters have [Z/H]
between --1 and --2 and dominate the average stellar metallicity at $z
> 8$.  Afterwards the metallicity increases itself by a factor of 30
to $[\bar{Z}_2/H] = -1.5$ through self-enrichment from a starburst.
Because this halo is located in a large-scale overdensity, most of the
ejecta falls back into the halo after reaching distances up to 20
comoving kpc, in turn keeping the halo gas metallicity high because
the inflows are relatively metal-rich themselves.  After the $z \sim
8$ starburst, the average stellar metallicity follows the average gas
metallicity within 0.1 dex.

@@

\textit{Intense halo}.--- The progenitors host a total of 56 Pop III
stars with 21 producing PISNe.  The first forms in a $6 \times 10^5
\Ms$ halo at $z=19$.  Pop III stars form on a regular interval in the
halo's progenitors until $z=9$ when most of these halos enter the
metal-rich bubble surrounding the intense halo.  There are two stars
forming at $z=7.7$ with [Z/H] = --5.7, whose molecular clouds were
enriched by external blastwaves that only partially enriched the dense
regions.  This falls into the metallicity range where fragmentation
occurs only if dust cooling is efficient below \tento{-4} $Z_\odot$.
The most massive progenitors of the intense halo cease to form Pop III
stars at $z<12$ because of nearby or intrinsic PISNe.

@@

\textit{Quiet halo}.--- A nearby PISNe provides a metallicity floor of
[Z/H] = --2.8 at which metallicity the first Pop II stars form.  The
stellar metallicity evolution exhibits what is expected from an
isolated system with the stellar feedback steadily enriching the ISM,
resulting in a correlation between stellar age and metallicity.  After
$z=10$ the metallicities plateau at [Z/H] = --2.1 for reasons
discussed in \S\ref{sec:halo}.  The SFR peaks at $z=10$ and decreases
as the cold gas reservoir is depleted.  Around $z=7.5$, a 25:1 minor
merger occurs, and the gas inside the satellite halo is compressed,
triggering metal-poor, [Z/H] = --3.2, SF during its nearest approach.
This halo infalls through a filament, where most of metal enrichment
in the quiet halo occurs in bi-polar flows perpendicular to the galaxy
disk and filament.  The filament and thus this satellite halo remains
relatively metal-poor.  Stars with [Z/H] $<$ --3 compose 1.6 percent
of the total stellar mass.

\textit{Intense halo}.--- In contrast with the quiet halo, the intense
halo undergoes a few mergers of halos with an established stellar
population.  This creates a superposition of age-metallicity tracks in
the SFH, seen in the complexity of Figure \ref{fig:pops}.  The first
two Pop II stellar clusters have an unusually high metallicity [Z/H]
$\sim$ --1, which occurs when a PISN blastwave triggers SF in two
neighboring halos.  Most of the early SF occurs around [Z/H] = --2.5.
At $z=9$, the halo's virial temperature reaches \tento{4} K.  This
combined with a 10:1 merger creates a starburst that quickly enriches
the halo to [Z/H] = --1.5 by $z=8$.  The halo continues to enrich
itself with further SF until $z=7$.  The spikes in the scatter plot
correspond to SN triggered SF in nearby molecular clouds
that are enriched up to a factor of 10 with respect to the ISM.
However the mass fraction of these metal-rich, SN triggered clusters
are small compared to the total stellar mass.  The starburst at $z=9$
creates a bimodal metallicity distribution with peaks at [Z/H] = --2.4
and --1.2 with the metal-rich component mainly being created after the
starburst.  Two systems with sizable stellar components merge into the
halo at $z \sim 8$, and their stellar populations are still
discernible in the metallicity-age plot, whose tracks are noted by
dashed lines.  Stars with [Z/H] $<$ --3 compose 1.8 percent of the
total stellar mass.

@@

We find that one PISN is sufficient to enrich the star-forming halo
and surrounding $\sim 5$ kpc to a metallicity of 10$^{-3} Z_\odot$,
given a Pop III characteristic mass of 100~\Ms.  If the first stars
have a lower characteristic mass that favor hypernovae
\cite{Tumlinson07_IMF}, then this metallicity floor should be lowered
by a factor of $\sim 10$ because (1) the metal ejecta is lowered by a
factor of $\sim 50$ and (2) the mixing mass is approximately decreased
by a factor of $(E_{\rm hyp}/E_{\rm PISN})^{3/5}$ in the Sedov-Taylor
solution, where $(E_{\rm hyp}/E_{\rm PISN}) \sim 0.1$ is the ratio of
explosion energies of a hypernova and PISN.  In the case where this
metallicity floor is less than the critical metallicity, then the next
instance of SF will further enrich the ISM, solidifying the transition
to Pop II SF.  We conclude that it only takes one, at most two, SNe
from Pop III stars in the halo progenitors to complete the transition
to Population II \cite{Frebel10}.  The question of
whether the critical metallicity is \tento{-6} or \tento{-3.5}
$Z_\odot$ is most applicable to nearby halos where the heavy elements
mix slowly into dense cores as the blastwave overtakes it.  Less than
two percent of stars have [Z/H] $< -3$ in both systems, consistent
with observations of metal-poor stars in the halo and dSphs
\cite{Beers05, Battaglia10}.

@@

\clearpage
\begin{figure*}
\includegraphics[width=0.88\textwidth]{f3.eps}
  \caption{\label{fig:projections} Density-weighted projections of gas
    density (top), temperature (middle), and metallicity (bottom) at
    $z=7$.  The left column shows the entire simulation volume, where
    the center and right columns focus on the intense and quiet halos,
    which are marked by arrows in the upper-left panel.  The
    metallicity projections are a composite picture of metals
    originating from Pop III (red) and Pop II (blue) stars with purple
    indicating a mixture of the two.}
\end{figure*}

@@

\clearpage
\begin{figure*}
% Needs f4b.eps
\includegraphics[width=0.88\textwidth]{f4a.eps}
  \caption{\label{fig:pops} The scatter plots show the SF history of
    the quiet (left) and intense (right) halos as a function of
    metallicity at $z=7$.  Each circle represents a star cluster,
    whose area is proportional to its mass.  The open circles in the
    upper right represent sizes of $10^3$, $10^4$, and $10^5$ \Ms~star
    clusters.  The dashed lines in the right panel guide the eye to
    two stellar populations that were formed in two satellite halos,
    merging at $z=7.5$.  The upper histogram collects the SF history
    into 25 equal temporal bins.  The right histogram depicts the
    stellar metallicity distribution.}
\end{figure*}

>>>>

The first (Pop III) stars are metal-free and have a large
characteristic mass and suppressed fragmentation in its protostellar
collapse \cite{ABN02, Bromm02_P3, Yoshida03, OShea07a}.  A fraction of
these stars enrich the surrounding intergalactic medium (IGM) when
they go supernova, which can happen in stars $\lsim 40~\Ms$ in Type II
supernovae (SNe) or in stars roughly between 140 \Ms~and 260 \Ms~in
pair-instability SNe (PISNe; \cite{2002ApJ...567..532H}).  The host
halo and the neighboring halos are then enriched with this ejecta.
There exists a critical metallicity that is $\sim 10^{-6} Z_\odot$ if
dust cooling is efficient \cite{Omukai05, Schneider06_Frag, clark08}
and $\sim 10^{-3.5} Z_\odot$ otherwise \cite{Bromm01,
  2009ApJ...691..441S}, where the gas can cool rapidly, lowering its
Jeans mass.  An intermediate characteristic mass of $\sim10~\Ms$ can
be occur if the gas cooling is suppressed to the cosmic microwave
background (CMB) temperature \cite{Larson98, Tumlinson07_IMF,
  2009ApJ...691..441S}.  The resulting Population II star cluster will
thus have a lower characteristic stellar mass than its metal-free
progenitors.  These first stellar clusters may be connected to stars
in the Milky Way halo and nearby dwarf spheroidal (dSph) galaxies,
both with a metallicity floor of [Z/H] = --4 \cite{Beers05,
  Tafelmeyer10, Frebel10_Obs}.

The transition from Pop III to Pop II star formation (SF) is solely
dependent on the propagation of metals from the SNe remnants into
future sites of SF.  Their flows are complex because of the
interactions between the SN blastwave, cosmological accretion and halo
mergers, and nearby stellar feedback.  In minihalos ($M \sim
10^6~\Ms$), radiation from a massive Pop III star can drive a 30
\kms~shock, which is 10 times greater than the escape velocity of the
halo, and leaves behind a warm ($3 \times 10^4$ K) and diffuse (0.5
\cubecm) medium \cite{Kitayama04, Whalen04, Abel07}.  This aids in the
expansion of the blastwave because it delays the transition to the
Sedov-Taylor and snowplow phases.  In PISNe, approximately half of the
metals stay in the IGM with a metal bubble size of a few kpc
\cite{Wise08_Gal, Greif10}.  The blastwave may induce SF in nearby
halos through the compression of the gas \cite{Ferrara98}, and
timescales for metal mixing into the dense gas are many dynamical
times \cite{Cen08} for shock velocities $\lsim100~\kms$.

Numerical simulations are useful to detangle and study these
complexities and the transition from Pop III to II stars.  In this
Report, we present the simulations that include both types of SF, and
their radiative and mechanical feedback.  The methods used here
incorporate and link together recent results from metal-enriched and
metal-free star formation, the critical metallicity, and
pair-instability supernovae.  This is the first simulation that can
resolve star-forming minihalos while following the transition to Pop
II star formation in the first galaxies.

@@

Here we present the gaseous and stellar evolution of two selected
halos in the simulation: one that has an early mass buildup but no
major mergers after $z=12$, and one that experiences a series of major
mergers between $z=10$ and $z=7$.  We name the halos ``quiet'' and
``intense'', respectively.

% We start with the mass accretion history and enrichment of the
% halos.  We then compare the nature of the stellar populations in
% these halos.

We illustrate the state of the simulation at $z=7$ in Figure
\ref{fig:projections} with density weighted projections of gas
density, temperature, and metallicity, showing the entire box and
focusing on the two halos of interest.  Radiative and mechanical
feedback create a multi-phase medium inside these halos, which are
embedded in a warm and ionized IGM.

@@

The gas fractions of both halos decrease from 0.15 to 0.08 by outflows
driven by ionization fronts and blastwaves in their initial
starbursts.  The quiet halo never recovers to the cosmic fraction
$\Omega_b/\Omega_M$ for the following reason.  The lack of major
mergers of halos above the filtering mass leads to a small gas
fraction because halos with $M < M_f$ are likely photo-evaporated,
hosting diffuse warm gas reservoirs instead of cold dense cores.
After $z=10$, the halo mainly accretes warm diffuse gas from the
filaments and IGM.  In contrast, the intense halo grows from major
mergers of halos with $M > M_f$.  The progenitor halos involved in the
major mergers are able to host molecular clouds and have, in general,
higher gas fractions.  Between $z=10$ and $z=8$, the gas fraction
increases from 0.07 to 0.12 until it jumps to 0.14 when a gas-rich
major merger occurs.  The stellar mass accordingly increases with the
ample supply of gas during this period.  The gas fraction then
decreases by 0.01 when galactic outflows are generated from the
starburst associated with this major merger.

\subsection*{Metallicity evolution}
\label{sec:zevo}

The evolution of the stellar and gas metallicity of both halos are
illustrated in Figure \ref{fig:evo}b.  PISNe from Pop III stars enrich
the nearby IGM out to a radius of 10 kpc and provides a metallicity
floor of $[Z_3/H] \sim -3$.  The metallicity from Pop II SNe initially
enrich the ISM of both halos to an average $[\bar{Z}_2/H]$ between
--3.5 and --3.

\textit{Quiet halo}.--- Afterwards an equilbrium of $[\bar{Z}_2/H]
\sim -2.5$ is established between metal-rich outflows and metal-poor
inflows.  Galactic outflows are directed in the polar directions of
the gas disk, keeping the adjacent filaments metal-poor.  These
features and a well-mixed ISM (cf.~\cite{Wise08_Gal, Greif10}) are
apparent in the metallicity projections in Figure
\ref{fig:projections}.  The average stellar metallicity is within 0.5
dex of $[\bar{Z}_2/H]$.

\textit{Intense halo}.--- The first few Pop II star clusters have
$[Z/H]$ between --1 and --2 and dominate the average stellar
metallicity at $z > 8$.  Afterwards the metallicity increases by a
factor of 30 to $[\bar{Z}_2/H] = -1.5$ through self-enrichment from a
starburst.  Because this halo is located in a large-scale overdensity,
most of the ejecta falls back into the halo after reaching distances
up to 20 comoving kpc, in turn keeping the halo gas metallicity high
because the inflows are relatively metal-rich themselves.  After the
$z \sim 8$ starburst, the average stellar metallicity follows the
average gas metallicity within 0.1 dex.

@@

\textit{Intense halo}.--- The progenitors host a total of 56 Pop III
stars with 21 producing PISNe.  The first forms in a $6 \times 10^5
\Ms$ halo at $z=19$.  Pop III stars form on a regular interval in the
halo's progenitors until $z=9$ when most of these halos enter the
metal-rich bubble surrounding the intense halo.  There are two stars
forming at $z=7.7$ with [Z/H] = --5.7, whose molecular clouds were
enriched by external blastwaves that only partially enriched the dense
regions.  This falls into the metallicity range where fragmentation
could occur if dust cooling is efficient below \tento{-4} $Z_\odot$.
The most massive progenitors of the intense halo cease to form Pop III
stars at $z<12$ because of nearby or intrinsic PISNe.

@@

\textit{Quiet halo}.--- A nearby PISNe provides a metallicity floor of
[Z/H] = --2.8 at which metallicity the first Pop II stars form.  The
stellar metallicity evolution exhibits what is expected from an
isolated system with the stellar feedback steadily enriching the ISM,
resulting in a correlation between stellar age and metallicity.  After
$z=10$ the metallicities plateau at [Z/H] = --2.1 for reasons
discussed in \S\ref{sec:zevo}.  The SFR peaks at $z=10$ and decreases
as the cold gas reservoir is depleted.  Around $z=7.5$, a 25:1 minor
merger occurs, and the gas inside the satellite halo is compressed,
triggering metal-poor, [Z/H] = --3.2, SF during its nearest approach.
This halo infalls through a filament, where most of metal enrichment
in the quiet halo occurs in bi-polar flows perpendicular to the galaxy
disk and filament.  The filament and thus this satellite halo remains
relatively metal-poor.  Stars with [Z/H] $<$ --3 compose 1.6 percent
of the total stellar mass.

\textit{Intense halo}.--- In contrast with the quiet halo, the intense
halo undergoes a few mergers of halos with an established stellar
population.  This creates a superposition of age-metallicity tracks in
the SFH, seen in the complexity of Figure \ref{fig:pops}.  The first
two Pop II stellar clusters have an unexpectedly high metallicity
[Z/H] $\sim$ --1, which occurs when a PISN blastwave triggers SF in
two neighboring halos.  Most of the early SF have [Z/H] = --2.5.  At
$z=9$, the halo's virial temperature reaches \tento{4} K.  This
combined with a 10:1 merger creates a starburst that quickly enriches
the halo to [Z/H] = --1.5 by $z=8$.  The halo continues to enrich
itself with further SF until $z=7$.  The spikes in the scatter plot
correspond to SN triggered SF in nearby molecular clouds that are
enriched up to a factor of 10 with respect to the ISM.  However their
mass fraction are small compared to the total stellar mass.  The
starburst at $z=9$ creates a bimodal metallicity distribution with
peaks at [Z/H] = --2.4 and --1.2 with the metal-rich component mainly
being created after the starburst.  Two systems with sizable stellar
components merge into the halo at $z \sim 8$, and their stellar
populations are still discernible in the metallicity-age plot, whose
tracks are noted by dashed lines.  Stars with [Z/H] $<$ --3 compose
1.8 percent of the total stellar mass.

@@

We find that one PISN is sufficient to enrich the star-forming halo
and surrounding $\sim 5$ kpc to a metallicity of 10$^{-3} Z_\odot$,
given $M_{\rm char} = 100~\Ms$.  If the first stars have a lower
characteristic mass that favor hypernovae \cite{Tumlinson07_IMF}, then
this metallicity floor should be lowered by a factor of $\sim 10$
because (1) the metal ejecta is lowered by a factor of $\sim 50$ and
(2) the mixing mass is approximately decreased by a factor of $(E_{\rm
  hyp}/E_{\rm PISN})^{3/5}$ in the Sedov-Taylor solution, where
$(E_{\rm hyp}/E_{\rm PISN}) \sim 0.1$ is the ratio of explosion
energies of a hypernova and PISN.  In the case where this metallicity
floor is less than the critical metallicity, then the next instance of
SF will further enrich the ISM, solidifying the transition to Pop II
SF.  We conclude that it only takes one, at most two, SNe from Pop III
stars in the halo progenitors to complete the transition to Population
II \cite{Frebel10}.  The question of whether the critical metallicity
is \tento{-6} or \tento{-3.5} $Z_\odot$ is most applicable to nearby
halos where the heavy elements mix slowly into dense cores as the
blastwave overtakes it.  Less than two percent of stars have [Z/H] $<
-3$ in both systems, consistent with observations of metal-poor stars
in the halo and dSphs \cite{Beers05, Battaglia10}.

@@

\clearpage
\begin{figure*}
\includegraphics[width=0.88\textwidth]{f3.eps}
\caption{\label{fig:projections} Density-weighted projections of gas
  density (top), temperature (middle), and metallicity (bottom) at
  $z=7$.  The left column shows the entire simulation volume, where
  the center and right columns focus on the intense and quiet halos,
  which are marked by left and right arrows in the upper-left panel.
  The metallicity projections are a composite picture of metals
  originating from Pop III (red) and Pop II (blue) stars with magenta
  indicating a mixture of the two.}
\end{figure*}

@@

\clearpage
\begin{figure*}
% Needs f4b.eps
\includegraphics[width=0.88\textwidth]{f4a.eps}
\caption{\label{fig:pops} The scatter plots show the SF history of the
  quiet (left) and intense (right) halos as a function of metallicity
  at $z=7$.  Each circle represents a star cluster, whose area is
  proportional to its mass.  The open circles in the upper right
  represent sizes of $10^3$, $10^4$, and $10^5$ \Ms~star clusters.
  The dashed lines in the right panel guide the eye to two stellar
  populations that were formed in two satellite halos, merging at
  $z=7.5$.  The upper histogram shows the SF rate.  The right
  histogram depicts the stellar metallicity distribution.}
\end{figure*}

### 2010-10-20 science_som.tex
Revised version after another reading.

In this section, we first describe our simulation setup.  We then
detail the SF models used here and how we model radiative and SNe
feedback.

@@

We use a non-equilibrium chemistry solver with 9 species of hydrogen,
helium, and molecular hydrogen \cite{Abel97}.  We will follow-up this
study with one that considers radiative cooling from metal-enriched
gas, using the method of \cite{2008MNRAS.385.1443S}.

@@

We distinguish Pop II and Pop III SF by the metallicity of the densest
cell in the molecular cloud.  Pop II stars are formed if [Z/H] $> -4$,
and Pop III stars are formed otherwise.  We do not consider Pop III.2
stars and intermediate mass stars from CMB-limited cooling.

@@

For Pop III stars, we use the same SF model as \cite{Abel07} and
\cite{Wise08_Gal} where each star particle represents a single star.
The critical overdensity in which a star forms is $5 \times 10^5$.
Instead of using a fixed stellar mass, we randomly sampled from an IMF
with a functional form of
%
\begin{equation}
\label{eqn:imf}
f(M)dM = M^{-1.3} \exp\left[-\left(\frac{M_{\rm char}}{M}\right)^{1.6}\right]
\end{equation}
to determine the stellar mass.  Above $M_{\rm char}$, it behaves as a
Salpeter IMF but is exponentially cutoff below that mass
\cite{Chabrier03, Clark09}.  For reproducibility, we record the
number of times the random number generator (Mersenne twister; \cite{MTwister})
has been called for use when restarting the simulations.

@@

We treat Pop II SF with the same prescription as \cite{Wise09}, which
is a modified version of the \cite{Cen92} method but accounts for the
fact that the molecular clouds are resolved.  The critical overdensity
is the same as the Pop III SF model.  In each star-forming region,
seven percent of the cold gas ($T < 10^3$ K) is removed from the grid
and deposited into the star particle.  Each star particle lives for 20
Myr, the maximum lifetime of an OB star.  These stars generate the
majority of the ionizing radiation and SNe feedback in stellar
clusters, thus we ignore lower mass stars.

@@

The mass-dependent luminosities and lifetimes of the Pop III stars are
taken from \cite{Schaerer02}.  The radiation field is evolved with
adaptive ray tracing \cite{Abel02_RT, Wise10} and is coupled
self-consistently to the hydrodynamics.  We model the \hh~dissociating
radiation with an optically-thin, inverse square profile, centered on
all stars.  These stars die as pair-instability SNe (PISNe) if they
are in the mass range 140--260 \Ms~\cite{Heger03}.  We use the
explosion energy from \cite{Heger02}, where we fit the following
function to their models, $E_{\rm PISN} = 10^{51} \times [5.0 + 1.304
(M_{\rm He} - 64)]$, where $M_{\rm He} = (13/24) \times (M_\star - 20)
\Ms$ is the helium core mass and $M_\star$ is the stellar mass.

The Pop II stars emit 6000 hydrogen ionizing photons per baryon over
their lifetime, and we do not consider singly- and doubly-ionizing
helium photons.  We note that low-metallicity stars generate up to a
factor of four more ionizing photons than a solar metallicity
population \cite{Schaerer03} and might be underestimating the
radiative feedback.  Nonetheless this study provides an excellent
first insight in the transition to Pop II SF, as the metal enrichment
is the key ingredient here.  For SN feedback, these stars generate
$6.8 \times 10^{48}$ erg s$^{-1}$ $\Ms^{-1}$ of energy after living
for 4 Myr, which is injected into spheres with a radius of 10 pc.  If
the resolution of the grid is less than 10/3 pc, we deposit the energy
into a $3^3$ cube surrounding the star particle.

>>>>

Here, we first describe our simulation setup.  We then detail the SF
models used here and how we model radiative and SNe feedback.

@@

We use a non-equilibrium chemistry solver with 9 species of hydrogen,
helium, and molecular hydrogen \cite{Abel97}.  We spatially
distinguish metal enrichment from Pop II and Pop III stars.  We will
follow-up this study with one that considers radiative cooling from
metal-enriched gas, using the method of \cite{2008MNRAS.385.1443S}.

@@

We distinguish Pop II and Pop III SF by the total metallicity of the
densest cell in the molecular cloud.  Pop II stars are formed if [Z/H]
$> -4$, and Pop III stars are formed otherwise.  We do not consider
Pop III.2 stars and intermediate mass stars from CMB-limited cooling.

@@

For Pop III stars, we use the same SF model as \cite{Wise08_Gal} where
each star particle represents a single star, forming at an overdensity
of $5 \times 10^5$.  Instead of using a fixed stellar mass, we randomly
sampled from an IMF with a functional form of
%
\begin{equation}
\label{eqn:imf}
f(M)dM = M^{-1.3} \exp\left[-\left(\frac{M_{\rm char}}{M}\right)^{1.6}\right]
\end{equation}
to determine the stellar mass.  Above $M_{\rm char}$, it behaves as a
Salpeter IMF but is exponentially cutoff below that mass
\cite{Chabrier03, Clark09}.  For reproducibility, we record the number
of times the random number generator (Mersenne twister;
\cite{MTwister}) has been called for use when restarting the
simulations.

@@

We treat Pop II SF with the same prescription as \cite{Wise09}, which
is a modified version of the \cite{Cen92} method but accounts for the
fact that the molecular clouds are resolved.  The critical overdensity
is the same as the Pop III SF model.  In each star-forming region,
seven percent of the cold gas ($T < 10^3$ K) is removed from the grid
and deposited into the star particle that lives for 20 Myr, the
maximum lifetime of an OB star.  These stars generate the majority of
the ionizing radiation and SNe feedback in stellar clusters, thus we
ignore lower mass stars.

@@

We use mass-dependent luminosities and lifetimes of the Pop III stars
from \cite{Schaerer02}.  The radiation field is evolved with adaptive
ray tracing \cite{Abel02_RT, Wise10} and is coupled self-consistently
to the hydrodynamics.  We model the \hh~dissociating radiation with an
optically-thin, inverse square profile, centered on all stars.  These
stars die as pair-instability SNe (PISNe) if they are in the mass
range 140--260 \Ms~\cite{Heger03}.  We use the explosion energy from
\cite{Heger02}, where we fit the following function to their models,
$E_{\rm PISN} = 10^{51} \times [5.0 + 1.304 (M_{\rm He} - 64)]$, where
$M_{\rm He} = (13/24) \times (M_\star - 20) \Ms$ is the helium core
(and equivalently the ejecta) mass and $M_\star$ is the stellar mass.

The Pop II stars emit 6000 hydrogen ionizing photons per baryon over
their lifetime, and we do not consider singly- and doubly-ionizing
helium photons.  We note that low-metallicity stars generate up to a
factor of four more ionizing photons than a solar metallicity
population \cite{Schaerer03} and might be underestimating the
radiative feedback.  Nonetheless this study provides an excellent
first insight in the transition to Pop II SF, as the metal enrichment
is the key ingredient here.  For SN feedback, these stars generate
$6.8 \times 10^{48}$ erg s$^{-1}$ $\Ms^{-1}$ after living for 4 Myr,
which is injected into spheres with a radius of 10 pc.  If the
resolution of the grid is less than 10/3 pc, we deposit the energy
into a $3^3$ cube surrounding the star particle.
